---
title: "User Excercise Classification"
author: "Hans Wu"
date: "April 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caret)
library(dplyr)
library(parallel)
library(doParallel)
```

## Executive Summary

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. This analysis is to classify 5 different correct and incorrect ways of performing barbell lifts from 6 participants based on accelerometers on the belt, forearm, arm, and dumbell.

## Prepare Data Set

### Load Data
```{r loadData}
# Load raw data
if (file.exists("./pml-training.csv")) {
  trainingRaw <- read.csv("./pml-training.csv")
} else {
  trainingRaw <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
}
if (file.exists("./pml-testing.csv")) {
  testingRaw <- read.csv("./pml-testing.csv")
} else {
  testingRaw <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))
}

head(trainingRaw)
```

### Clean NA values
From glancing through the data, there are some columns with mostly missing values. Will exclude those columns from the training data
Based on the experiments, the time related variables are not likely to affect the outcome, will exclude those as well.
```{r cleanNA}
# Check fo N/A values
naCount <- apply(trainingRaw, 2, function(x) sum(is.na(x)))

# Exclude columns that have majority values as missing
excludeNACols <- names(naCount[naCount > 0])

trainingNonNA <- trainingRaw %>%
  # Exclude outcome
  select(-classe) %>%
  select(-one_of(excludeNACols)) %>%
  # Exclude time related columns since it's mostly likely irrelevant
  select(-one_of(c("raw_timestamp_part_1", "raw_timestamp_part_2", "num_window", "new_window", "cvtd_timestamp"))) %>%
  # Exclude index column
  select(-X)
```

## Model Selection
### Explore Correlation between Continuous Variables
Check correlations between continuous variables and the variables with high correlations with the variables that has been seen.
```{r exploreContVars}
# Get continous variables
numericCol <- names(trainingNonNA[,sapply(trainingNonNA[1,], is.numeric)])
numericTraining <- trainingNonNA %>% select(one_of(numericCol))
# Generate correlation matrix
corTraining <- cor(numericTraining)

highCorCol <- rep(FALSE, length(numericCol))
rowEnd <- length(highCorCol)
# Loop through the bottom left half of the matrix
for (i in 1:(rowEnd - 1)) {
  startingRow <- i + 1
  col <- i
  # captures high correlated variables
  highCorCol[startingRow:rowEnd] <- highCorCol[startingRow:rowEnd] | (abs(corTraining[startingRow:rowEnd,col]) >= 0.75)
}
# Include select variables more likely to be independent
lowCorColName <- names(numericTraining)[-highCorCol]
```

### Explore Categorical Variables
```{r exploreCategVars}
categoricalTraining <- trainingNonNA %>%
  select(-one_of(c(numericCol, "user_name")))

sum(categoricalTraining == "" | categoricalTraining == "#DIV/0!")

chisq.test(trainingNonNA$user_name, trainingRaw$classe)$p.value
```
Besides user_name, most of the categorical variables seems to be continuous variables, but the majority is unknown value such as empty string and "#DIV/0!". Therefore, these categorical variables will not be selected for the model.
Chisq test between user_name and classe gives a very low p-value (3.67e-40) which indicates user_name has relation with classe.

## Model Training
Will train the model using randome forest with 10-fold cross validation.
```{r trainModel}
finalTraining <- trainingNonNA %>%
  select(one_of(c(lowCorColName, "user_name")))
finalTraining$classe <- trainingRaw$classe

set.seed(7)
# Use 10 fold cross validation
trainCtrl <- trainControl(method="cv", number=10, savePredictions=TRUE, allowParallel = TRUE)

cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
model <- train(classe ~ ., method="rf", data=finalTraining, trControl=trainCtrl)

stopCluster(cluster)
registerDoSEQ()

model
```
The model accuracy is abount 99.5%.

## Prediction
Predict outcome of the testing data set.
```{r prediction}
predict(model, testingRaw)
```